---
title: "RDataWrangling"
author: "Çağrı Çebişli"
date: "2023-01-22"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#   Useful Links

https://www.uvm.edu/~tdonovan/RforFledglings/data-wrangling-with-dplyr.html

https://intro2r.com/summarising-data-frames.html

#   For Importing Data

##    Folders, Working Directory, Shortcuts

-   Shortcut for chunk: Command + Option + I

-   Invisible Chunk: r,warning=FALSE,error=FALSE,message=FALSE,include=FALSE, Echo=FALSE

**Sets Working Directory to Saved Folder:**

setwd(dirname(rstudioapi::getActiveDocumentContext()$path))

library(here): read_xlsx(here("folder name", "data to import.xlsx"))

After you set your working directory to saved folder, you can move around with using here package. Thus, this operation is good for conducting reproducible research. You do not talk about /c:cagri etc., only move around folders, so folder name is important here.

##    Exploratory Data Analysis

**EDA Report DlookR**

library(dlookr)

EDA Report

data %>% eda_report(output_format = "html", output_file   = "EDA_diamonds.html")

Very good exploratory data analysis page. Do take notes while reading this, write down ideas and disaggregations.

###   Crosstabs

Good crosstab to see frequency, include NA to understand overall column.

**Some x-tab examples**

(table(data$column,useNA = "always"))

prop.table(table(data$column,useNA="always"))

CD_WFP %>% group_by(province) %>% summarize(AvgTR_Month = mean(Arrival_Differences,na.rm = T),
                                            CampRes = mean(camp_residence_how_long,na.rm = T))


***
aggregate(flowers[, 4:7], by = list(nitrogen = flowers$nitrogen,
                                 treat = flowers$treat), FUN = mean)
***



#   Column Re-arrange

Extract month from character saved date variable.

Dots in format can change into: % , / , etc.

month(as.POSIXlt(data$date_as_character_variable, format="%d.%m.%Y"))

#   Select Columns - Arrange - Filter

```{r}
library(tidyr)
library(dplyr)
```

```{r}
starwars %>%
  mutate(name, bmi = mass / ((height / 100)  ^ 2)) %>%
  select(name:mass, bmi)
```

```{r}
starwars %>%
  arrange(desc(mass))
```

```{r}
starwars %>%
  group_by(species) %>%
  summarise(
    n = n(),
    mass = mean(mass, na.rm = TRUE)
  ) %>%
  filter(
    n > 1,
    mass > 50
  )
```

##    Subset

subset(flowers, treat == "tip" & nitrogen == "medium" & block == 2, 
                            select = c("treat", "nitrogen", "leafarea"))
                            
##    Reshape- Long and Wide Data

###   Wide Data into Long Data

```{r}
wide_data <- data.frame(subject = c("A", "B", "C", "D"),
               sex = c("M", "F", "F", "M"),
               control = c(12.9, 5.2, 8.9, 10.5),
               cond1 = c(14.2, 12.6, 12.1, 12.9),
               cond2 = c(8.7, 10.1, 14.2, 11.9))
wide_data
```


```{r}
library(reshape2)

my_long_df <- melt(data = wide_data, id.vars = c("subject", "sex"),
                   measured.vars = c("control", "cond1", "cond2"),
                   variable.name = "condition", value.name = "measurement")
my_long_df
```

###   Long Data into Wide Data

```{r}
long_data <- data.frame(
             subject = rep(c("A", "B", "C", "D"), each = 3),
             sex = rep(c("M", "F", "F", "M"), each =3),
             condition = rep(c("control", "cond1", "cond2"), times = 4),
             measurement = c(12.9, 14.2, 8.7, 5.2, 12.6, 10.1, 8.9,
                             12.1, 14.2, 10.5, 12.9, 11.9))
head(long_data,5)
```


```{r}
my_wide_df <- dcast(data = long_data, subject + sex ~ condition, 
                    value.var = "measurement")
my_wide_df
```


```{r}


```



```{r}

```


```{r}
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```

